### 2015-11-26 Irregular-occ/Irregular-occ.tex
.

\title{Stellar occultation predictions for  10 irregular satellites of giant planets and Triton for 2015-2017
%\fnmsep\thanks{Table 8 are only available in electronic form at the CDS via anonymous ftp to cdsarc.u-strasbg.fr (130.79.128.5) or via http://cdsweb.u-strasbg.fr/cgi-bin/qcat?J/A+A/ and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{The complete version of Table 8 is available through CDS and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{Partially based on observations made at Laborat\'orio Nacional de Astrof\'{\i}sica (LNA), Itajub\'a-MG, Brazil.}
%\fnmsep\thanks{Partially based on observations through the ESO runs 079.A-9202(A), 075.C-0154, 077.C-0283 and 079.C-0345.}
%\fnmsep\thanks{\textbf{Partially based on observations made at Observatoire de Haute Provence (OHP), F-04870 Saint-Michel l'observatoire, France}}
}

\author{ A. R. Gomes-Júnior
          \inst{1},
          M. Assafin 
          \inst{1} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France},
          L. Beauvalet
          \inst{2,3},
          R. Vieira-Martins
          \inst{1,2,4} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France},
%          J.-E. Arlot
%          \inst{4},      
          J. I. B. Camargo
          \inst{2,4},     
          B. E. Morgado
          \inst{1,2}
%          F. Braga-Ribas
%          \inst{2,4}, 
%          D. N. da Silva Neto
%          \inst{6},
%          A. H. Andrei
%          \inst{1,2} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau, 75014 Paris, France},
%          A. Dias-Oliveira
%          \inst{2},
%          G. Benedetti-Rossi
%          \inst{2},
%          Y. Duchemin
%          \inst{4,7},
%          J. Desmars
%          \inst{4},
%          V. Lainey
%          \inst{4},
%          W. Thuillot
%          \inst{4}
          }

@@

   \institute{Observatório do Valongo/UFRJ, Ladeira Pedro Antônio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
              \email{altair08@astro.ufrj.br}
              \and
Observatório Nacional/MCT, R. General José Cristino 77,
CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
              \email{rvm@on.br}
              \and
Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France
              \and
Laboratório Interinstitucional de e-Astronomia - LIneA, Rua Gal. José Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil
%              \and
%Institut de mécanique céleste et de calcul des éphémérides - Observatoire %de Paris, UMR 8028 du CNRS,
%77 Av. Denfert-Rochereau, 75014 Paris, France\\
%              \email{arlot@imcce.fr}
%              \and
%Federal University of Technology - Paraná (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
%              \and
%Centro Universitário Estadual da Zona Oeste, Av. Manual Caldeira de Alvarenga 1203, CEP 23.070-200 Rio de Janeiro RJ, Brazil
%              \and
%              ESIGELEC-IRSEEM, Technopôle du Madrillet, Avenue Galilée, 76801 Saint-Etienne du Rouvray, France
              }

@@

\abstract
{Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, albedo and composition, to trace back their true origin. The best ground-based technique to constrain those parameters is the observation of stellar occultations by these objects.}
{We aim to predict stellar occultations for the eight biggest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda; the irregular satellite of Saturn: Phoebe; and the satellites of Neptune: Nereid and Triton.}
{We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. We used a new ephemeris based solely on the observations from \cite{GomesJunior2015} to generate quick predictions for the short-time future for the satellites of Jupiter. For Phoebe, Nereid and Triton we used the JPL ephemeris.}
{We managed to identify 687 candidates of stellar occultations between the period of January, 2015 and December, 2017. We made observational tests for the predictions of the events of Himalia in March 03, 2015 and Elara in March 30, 2015. The stars and objects were observed close to the date predicted and in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions.}
{The comparisons between the predictions and the observation tests show a good agreement. We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.}

@@

\section{Introduction} \label{Sec: introducao}

Irregular satellites revolve around giant planets at large distances, on eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around of their planet, but were captured by their planets in the early solar system \citep{Sheppard2005}.

There are several of capture mechanisms of objects by giant planets suggested in the literature. For example the Gas Drag in the primordial circumplanetary nebulae \citep{Sheppard2005} where the object would be affected by the gas drag and its velocity slowed down until it be captured by the planet. Another mechanism is called pull-down capture \citep{Sheppard2005}, where the mass of the planet would increase while the object was temporarily captured.

A mechanism based in the Nice model \citep{Morbidelli2005, Tsiganis2005, Gomes2005} was proposed by \cite{Nesvorny2007} and, in the specific case of Jupiter with the modern Nice model, by \citealp{Nesvorny2014}. In this scenario, during the early solar system instability, encounters between the outer planets occurred. These planetary encounters could exchange energy and angular momentum between planets and the objects nearby allowing the capture of irregular bodies by the giant planets. In this scenario, the survival rate of prior-LHB (Late Heavy Bombardment) satellites is very small.

Another possible mechanism is the capture through collisional interactions \citep{Sheppard2005}. A collision between two small bodies in the Hill's sphere of the planet could generate fragmented objects and the dissipated energy could be such that some of these objects could be captured.

Some of these objects are in dynamical groups with similar semi-major axis, eccentricities and inclinations, called families, similar to families found in the Main Asteroid Belt. These families may have been created by a parent body disrupted by collisions with comets or other satellites \citep{Nesvorny2004}. These collisions are more likely to have occurred during the LHB \citep{Gomes2005}.

\cite{Nesvorny2003} studied the collision rates between irregular satellites and concluded that some satellites could have been removed by collision with a bigger satellite. The collision rate between satellites of the Himalia Group (Himalia, Elara, Lysithea and Leda, mainly), for instance, was found to be more than one during the solar system age suggesting that their current structure was originated by satellite-satellite collision.

For Phoebe, ejected material from its surface caused by impacts could evolve due to Poynting-Robertson drag and collide with Iapetus causing the large variation in albedo observed on it \citep{Nesvorny2003}. Indeed, Cassini detected in Phoebe an absorption feature at 2.42 $\mu m$ (probably CN combinations) that was also detected in the dark side of Iapetus \citep{Clark2005}.

The region of origin of these object is not well known. \cite{Grav2003} and \cite{Grav2007} showed that the irregular satellites from the giant planets have their colors and spectral slopes similar to C-, D- and P-type asteroids, Centaurs and trans-neptunian objects (TNOs) suggesting that they come from different locations in the early solar system.

In order to obtain precise fundamental physical parameters like size, shape and albedo for the irregular satellites and thus to contribute to the study of their origin, we want to observe stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

No observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have prevented earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are already good enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, the cumulated uncertainty of ephemeris and star position must be around 40 mas to get good chances of observing a stellar occultation, which is quite feasible today.

From what we have just presented, we can think that the observation of an stellar occultation by an irregular satellite should be likelier than occultations by TNOs. The orbits of their host planets are well known and their observation time-span covers many orbital periods, contrary to TNOs.
%Unlike stellar occultations by TNOs, which nevertheless have been proved effective, the observation of an occultation by a irregular satellite is, in principle, more favorable. The orbits of their host planets are well known and these satellites have been observed completing already many orbital periods around them, thus presenting better ephemeris than TNOs. 
Moreover, the irregular satellites are closer to Earth which means a minor localization error in kilometer.

\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites. For an object at the distance of Jupiter, this represents an error bigger than 700 km in the shadow path. Using the positions obtained by \cite{GomesJunior2015} we produced a specific ephemeris for short-time future for the satellites of Jupiter and better predict stellar occultations for these objects.

We present in this paper stellar occultation predictions for the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda), Phoebe from Saturn and Nereid and Triton from Neptune. In the section \ref{Sec: Rationale} we explore the scientific rationale for the study of the irregular satellites and the possibility of having a common origin with TNOs. In section \ref{Sec: integration} we show the the process of the production of the new ephemeris. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites and how they were made. Some tests made to confirm the predictions are presented in section \ref{Sec: testes} and the final discussion are in section \ref{Sec: discussion}.

\section{Scientific Rationale} \label{Sec: Rationale}

%\textcolor{red}{Como pode ver, comparado ao paragrafo na Introducao que falou a respeito, o texto aqui está muito pequeno e chovendo no molhado.}

%\textcolor{red}{Tem que falar mais. Citar o paper no livro da Barucci que voce me mostrou outro dia. Dar mais citacoes sobre essa hipotese. Falar dos albedos parecidos, cores. Tem que fundamentar a hipotese de irregulares = TNOs pequenos.}

%\textcolor{red}{Tambem falar aqui das vantagens de se observar ocultacoes dos irregulares, porque apesar de alguns satelites terem efemerides ruins, outros nem tanto, as efemerides sao melhores que a de TNOs (os irregulares ja completaram voltas em torno dos seus planetas); no caso de Jupiter, ele esta proximo o que ajuda a diminuir o erro em km.}

%\textcolor{red}{E dizer que, mesmo que a hipotese de TNOs que nos motiva a fazer as predicoes, venha a ser refutada no futuro, seja por outros trabalhos, seja pelos resultados futuros de observacoes de nossas proprias ocultacoes, o levantamento em si das propriedades fisicas, tamanho, forma e albedo ainda assim seriam muito uteis para testar as hiposteses de captura, ou seja, contribui para o estudo da formacao e evolucao dos sistemas de Jupiter, Saturno e Netuno.}

There is no consensus for a single model explaining where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved having their physical properties similar to those they had when they were formed \citep{Camargo2014}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. A clever way to overcome this difficulty is to study  the much closer irregular satellites, under the hypothesis that they share a common origin with the small TNOs' population.

Triton is a uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. Differently, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007}.

@@

\section*{Introduction V2}

Irregular satellites revolve around giant planets at large distances, on eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around of their planet, but were captured by their planets in the early solar system \citep{Sheppard2005}.

There is no consensus for a single model explaining where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

@@

\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved having their physical properties similar to those they had when they were formed \citep{Camargo2014}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. A clever way to overcome this difficulty is to study  the much closer irregular satellites, under the hypothesis that they share a common origin with the small TNOs' population.

@@

In order to obtain precise fundamental physical parameters like size, shape and albedo for the irregular satellites and thus to contribute to the study of their origin, we want to observe stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

@@

Triton is a uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. Differently, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007}.

No observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have prevented earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are already good enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, the cumulated uncertainty of ephemeris and star position must be around 40 mas to get good chances of observing a stellar occultation, which is quite feasible today.

@@

The last observations used to develop JPL current ephemeris of the irregular satellites of Jupiter were obtained in 2012 \citep{Jacobson2012}. As a result, the errors on the residuals of recent observations are too important for the ephemeris to accurately predict stellar occultations without any corrections. In order to have efficient ephemerides for predicting the stellar occultations by irregular satellites, we need predictions based on recent observations. This is why we decided to develop our own ephemeris based on the observations published in \cite{GomesJunior2015}.

>>>>

\title{Stellar occultation predictions for  10 irregular satellites of giant planets plus Triton for 2015-2017
%\fnmsep\thanks{Table 8 are only available in electronic form at the CDS via anonymous ftp to cdsarc.u-strasbg.fr (130.79.128.5) or via http://cdsweb.u-strasbg.fr/cgi-bin/qcat?J/A+A/ and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{The complete version of Table 8 is available through CDS and IAU NSDC data base at www.imcce.fr/nsdc.}
%\fnmsep\thanks{Partially based on observations made at Laborat\'orio Nacional de Astrof\'{\i}sica (LNA), Itajub\'a-MG, Brazil.}
%\fnmsep\thanks{Partially based on observations through the ESO runs 079.A-9202(A), 075.C-0154, 077.C-0283 and 079.C-0345.}
%\fnmsep\thanks{\textbf{Partially based on observations made at Observatoire de Haute Provence (OHP), F-04870 Saint-Michel l'observatoire, France}}
}

\author{ A. R. Gomes-Júnior
          \inst{1},
          M. Assafin 
          \inst{1} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France},
          L. Beauvalet
          \inst{2,3},
          R. Vieira-Martins
          \inst{1,2,4} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France},
%          J.-E. Arlot
%          \inst{4},      
          J. I. B. Camargo
          \inst{2,4},     
          B. E. Morgado
          \inst{1,2}
          F. Braga-Ribas
          \inst{2,5}, 
%          D. N. da Silva Neto
%          \inst{6},
%          A. H. Andrei
%          \inst{1,2} \fnmsep\thanks{Affiliated researcher at Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau, 75014 Paris, France},
%          A. Dias-Oliveira
%          \inst{2},
%          G. Benedetti-Rossi
%          \inst{2},
%          Y. Duchemin
%          \inst{4,7},
%          J. Desmars
%          \inst{4},
%          V. Lainey
%          \inst{4},
%          W. Thuillot
%          \inst{4}
          }

@@

   \institute{Observatório do Valongo/UFRJ, Ladeira Pedro Antônio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
              \email{altair08@astro.ufrj.br}
              \and
Observatório Nacional/MCT, R. General José Cristino 77,
CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
              \email{rvm@on.br}
              \and
Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France
              \and
Laboratório Interinstitucional de e-Astronomia - LIneA, Rua Gal. José Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil
%              \and
%Institut de mécanique céleste et de calcul des éphémérides - Observatoire %de Paris, UMR 8028 du CNRS,
%77 Av. Denfert-Rochereau, 75014 Paris, France\\
%              \email{arlot@imcce.fr}
              \and
Federal University of Technology - Paraná (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
%              \and
%Centro Universitário Estadual da Zona Oeste, Av. Manual Caldeira de Alvarenga 1203, CEP 23.070-200 Rio de Janeiro RJ, Brazil
%              \and
%              ESIGELEC-IRSEEM, Technopôle du Madrillet, Avenue Galilée, 76801 Saint-Etienne du Rouvray, France
              }

@@

\abstract
{Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects.}
{We aim to predict stellar occultations for the eight biggest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda, and for the irregular satellites Phoebe of Saturn and Nereid of Neptune, and also for Triton.}
{We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. We used a new ephemeris based solely on the observations from \cite{GomesJunior2015} to generate predictions for the short-time future for the satellites of Jupiter. For Phoebe, Nereid and Triton we used the JPL ephemeris.}
{We managed to identify 687 candidates of stellar occultations between the period of January, 2015 and December, 2017. We made observational tests for the prediction of the event of Himalia in March 03, 2015. The stars and objects were observed close to the date predicted and in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions.}
{The comparisons between the predictions and the observation tests show a good agreement. We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.}

@@

%\section{Introduction} \label{Sec: introducao} 
%
%Irregular satellites revolve around giant planets at large distances, on eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around of their planet, but were captured by their planets in the early solar system \citep{Sheppard2005}.
%
%There are several of capture mechanisms of objects by giant planets suggested in the literature. For example the Gas Drag in the primordial circumplanetary nebulae \citep{Sheppard2005} where the object would be affected by the gas drag and its velocity slowed down until it be captured by the planet. Another mechanism is called pull-down capture \citep{Sheppard2005}, where the mass of the planet would increase while the object was temporarily captured. 
%
%A mechanism based in the Nice model \citep{Morbidelli2005, Tsiganis2005, Gomes2005} was proposed by \cite{Nesvorny2007} and, in the specific case of Jupiter with the modern Nice model, by \citealp{Nesvorny2014}. In this scenario, during the early solar system instability, encounters between the outer planets occurred. These planetary encounters could exchange energy and angular momentum between planets and the objects nearby allowing the capture of irregular bodies by the giant planets. In this scenario, the survival rate of prior-LHB (Late Heavy Bombardment) satellites is very small.
%
%Another possible mechanism is the capture through collisional interactions \citep{Sheppard2005}. A collision between two small bodies in the Hill's sphere of the planet could generate fragmented objects and the dissipated energy could be such that some of these objects could be captured.
%
%Some of these objects are in dynamical groups with similar semi-major axis, eccentricities and inclinations, called families, similar to families found in the Main Asteroid Belt. These families may have been created by a parent body disrupted by collisions with comets or other satellites \citep{Nesvorny2004}. These collisions are more likely to have occurred during the LHB \citep{Gomes2005}.
%
%\cite{Nesvorny2003} studied the collision rates between irregular satellites and concluded that some satellites could have been removed by collision with a bigger satellite. The collision rate between satellites of the Himalia Group (Himalia, Elara, Lysithea and Leda, mainly), for instance, was found to be more than one during the solar system age suggesting that their current structure was originated by satellite-satellite collision.
%
%For Phoebe, ejected material from its surface caused by impacts could evolve due to Poynting-Robertson drag and collide with Iapetus causing the large variation in albedo observed on it \citep{Nesvorny2003}. Indeed, Cassini detected in Phoebe an absorption feature at 2.42 $\mu m$ (probably CN combinations) that was also detected in the dark side of Iapetus \citep{Clark2005}.
%
%The region of origin of these object is not well known. \cite{Grav2003} and \cite{Grav2007} showed that the irregular satellites from the giant planets have their colors and spectral slopes similar to C-, D- and P-type asteroids, Centaurs and trans-neptunian objects (TNOs) suggesting that they come from different locations in the early solar system.
%
%In order to obtain precise fundamental physical parameters like size, shape and albedo for the irregular satellites and thus to contribute to the study of their origin, we want to observe stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.
%
%No observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have prevented earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are already good enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, the cumulated uncertainty of ephemeris and star position must be around 40 mas to get good chances of observing a stellar occultation, which is quite feasible today.
%
%\begin{table}
%\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
%\begin{centering}
%\begin{tabular}{lccc}
%\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
%Satellite  & mas\tablefootmark{a}  & km & References \tabularnewline
%\hline
%Ananke & 8 & 29 & 1 \tabularnewline
%Carme & 13 & 46 & 1 \tabularnewline
%Elara & 24 & 86 & 1 \tabularnewline
%Himalia & 41 & $(150\times120) \pm 20$\tablefootmark{b} & 2 \tabularnewline
%Leda & 5 & 20 & 1 \tabularnewline
%Lysithea & 10 & 36 & 1 \tabularnewline
%Pasiphae & 17 & 62 & 1 \tabularnewline
%Sinope & 10 & 37 & 1 \tabularnewline
%\hdashline
%Phoebe & 32 & $212 \pm 1.4$\tablefootmark{b} & 3 \tabularnewline
%\hdashline
%Nereid & 15 & $340 \pm 50$\tablefootmark{c} & 4 \tabularnewline
%Triton & 124 & $2707 \pm 2.0$\tablefootmark{c} & 5 \tabularnewline
%\hline
%\end{tabular}
%\tablebib{
%(1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}; (4) \cite{Thomas1991}; (5) \cite{Thomas2000}.}
%\end{centering}
%\tablefoottext{a}{Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
%\tablefoottext{b}{From Cassini observation.}
%\tablefoottext{c}{From Voyager-2 observation.}
%\par
%\end{table}
%
%From what we have just presented, we can think that the observation of an stellar occultation by an irregular satellite should be likelier than occultations by TNOs. The orbits of their host planets are well known and their observation time-span covers many orbital periods, contrary to TNOs.
%%Unlike stellar occultations by TNOs, which nevertheless have been proved effective, the observation of an occultation by a irregular satellite is, in principle, more favorable. The orbits of their host planets are well known and these satellites have been observed completing already many orbital periods around them, thus presenting better ephemeris than TNOs. 
%Moreover, the irregular satellites are closer to Earth which means a minor localization error in kilometer.
%
%\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites. For an object at the distance of Jupiter, this represents an error bigger than 700 km in the shadow path. Using the positions obtained by \cite{GomesJunior2015} we produced a specific ephemeris for short-time future for the satellites of Jupiter and better predict stellar occultations for these objects.
%
%We present in this paper stellar occultation predictions for the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda), Phoebe from Saturn and Nereid and Triton from Neptune. In the section \ref{Sec: Rationale} we explore the scientific rationale for the study of the irregular satellites and the possibility of having a common origin with TNOs. In section \ref{Sec: integration} we show the the process of the production of the new ephemeris. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites and how they were made. Some tests made to confirm the predictions are presented in section \ref{Sec: testes} and the final discussion are in section \ref{Sec: discussion}.
%
%\section{Scientific Rationale} \label{Sec: Rationale}
%
%%\textcolor{red}{Como pode ver, comparado ao paragrafo na Introducao que falou a respeito, o texto aqui está muito pequeno e chovendo no molhado.}
%
%%\textcolor{red}{Tem que falar mais. Citar o paper no livro da Barucci que voce me mostrou outro dia. Dar mais citacoes sobre essa hipotese. Falar dos albedos parecidos, cores. Tem que fundamentar a hipotese de irregulares = TNOs pequenos.}
%
%%\textcolor{red}{Tambem falar aqui das vantagens de se observar ocultacoes dos irregulares, porque apesar de alguns satelites terem efemerides ruins, outros nem tanto, as efemerides sao melhores que a de TNOs (os irregulares ja completaram voltas em torno dos seus planetas); no caso de Jupiter, ele esta proximo o que ajuda a diminuir o erro em km.}
%
%
%%\textcolor{red}{E dizer que, mesmo que a hipotese de TNOs que nos motiva a fazer as predicoes, venha a ser refutada no futuro, seja por outros trabalhos, seja pelos resultados futuros de observacoes de nossas proprias ocultacoes, o levantamento em si das propriedades fisicas, tamanho, forma e albedo ainda assim seriam muito uteis para testar as hiposteses de captura, ou seja, contribui para o estudo da formacao e evolucao dos sistemas de Jupiter, Saturno e Netuno.}
%
%There is no consensus for a single model explaining where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.
%
%Phoebe is the most studied irregular satellite. \cite{Clark2005} suggest that its surface is probably covered by material of cometary origin. It was also stated by \cite{Johnson2005} that if the porosity of Phoebe is 15\%, Phoebe would have an uncompressed density similar to those of Pluto and Triton.
%
%\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved having their physical properties similar to those they had when they were formed \citep{Camargo2014}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. A clever way to overcome this difficulty is to study  the much closer irregular satellites, under the hypothesis that they share a common origin with the small TNOs' population.
%
%Triton is a uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}. 
%
%Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. Differently, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007}.

@@

\section{Introduction}\label{Sec: introducao}

Irregular satellites revolve around giant planets at large distances on eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around their planet, but were captured in the early solar system \citep{Sheppard2005}. There is no consensus for a single model explaining where the irregular satellites were formed.

\cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

@@

\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or TNOs. TNOs are highly interesting objects that, due to their large heliocentric distances, may be highly preserved and have physical properties similar to those they had when they were formed \citep{Barucci2008}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. Thus, if irregular satellites - or at least a few of them - do share a common origin with small TNOs, and since these objects are situated at much closer distances, this gives a unique chance of observing and studying representatives of this specific TNO population in much greater detail than could ever be possible by direct observation of this population in the Kuiper Belt.

@@

In order to obtain precise fundamental physical parameters like size and shape, thus constraining the albedo and in a broader sense also the composition for the irregular satellites and therefore to contribute to the study of their origin, we aim at observing stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

@@

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}.

No observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are precise enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, the cumulated uncertainty of ephemeris and star position must be around 40 mas to get good chances of observing a stellar occultation, which is quite feasible today.

@@

The last observations used to develop JPL current ephemeris of the irregular satellites of Jupiter were obtained in 2012 \citep{Jacobson2012}. As a result, the errors in the JPL ephemeris for the current epoch are large enough to prevent accurate predictions of stellar occultations without any corrections. %In order to have efficient ephemerides for predicting the stellar occultations by irregular satellites, we need predictions based on recent observations. This is why we decided to develop our own ephemeris based on the observations published in \cite{GomesJunior2015}.
